---
title: "R Project Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Report storyline

#### Chapter 1
- Problem understanding = Introduction on poverty
- Research question formulation:

Main question: 
How did poverty develop around the world since 1990 to national and absolute dimension and how is this related to the presence of social safety programs?

To answer this research, several sub questions are defined:

-  Question 1: How did poverty for all develop from 1990 – 2015 for national and absolute dimensions?
	- To national definitions  world map
	- To UN definitions 1.90,350,5,50 -> stacked linechart 
	- Distribution for men/women/children stacked bar chart for three groups
- Question 2: How is wealth distributed for countries with a relatively high poverty headcount to national poverty lines.
	- GINI index and Lorenz curve for pick five worst from national map
- Question 3: What is the effect of social programs on the prevalence of poverty according to national poverty lines?
	- Barchart of absolute and percentage population living under national poverty line
	- Regression with seaborn (python example) x = poverty, y social program
	
#### Chapter 2

- Data understanding  introduce used databases + data descriptives
- Data preparation  how did we modify the data to properly use it 
#### Chapter 3 modelling and graphs
-	Define a descriptive model for the relevant data (Visual graphs)
-	Specify a prior distribution
-	Use Bayesian inference to discuss credibility across parameter values
#### Chapter 4 evaluation
- evaluation
-	Discuss research findings to quality of the model, quality of the data. 
-	How could the model be improved  suggest next steps
#### Chapter 5 development
	- Can we predict future behaviour/Guide additional policy
	- How would we communicate our findings to decision makers? 



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
