---
title: "R Project Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Report storyline

#### Chapter 1
- Problem understanding = Introduction on poverty
- Research question formulation:

Main question: 
How did poverty develop around the world since 1990 to national and absolute dimension and how is this related to the presence of social safety programs?

To answer this research, several sub questions are defined:

-  Question 1: How did poverty for all develop from 1990 – 2015 for national and absolute dimensions?
	- To national definitions  world map
	- To UN definitions 1.90,350,5,50 -> stacked linechart 
	- Distribution for men/women/children stacked bar chart for three groups
- Question 2: How is wealth distributed for countries with a relatively high poverty headcount to national poverty lines.
	- GINI index and Lorenz curve for pick five worst from national map
- Question 3: What is the effect of social programs on the prevalence of poverty according to national poverty lines?
	- Barchart of absolute and percentage population living under national poverty line
	- Regression with seaborn (python example) x = poverty, y social program
	
#### Chapter 2

- Data understanding  introduce used databases + data descriptives

Our data comes from World Development Indicators database(http://data.worldbank.org/data-catalog/world-development-indicators). We focus on one of the UN Sustainable Development Goals, No Poverty. Data of over 1440 indicators of more than 200 countries and regions from 1960 to 2015 are provided in the WDI dataset, but only some of them are relevant to what we want to discuss about.  
First, we narrow down the scope of indicators of the following fields that might be related to ‘No poverty’ goal:

1. Poverty headcount and poverty gap:
-Poverty headcount ratio at national poverty lines (% of population)
-Rural poverty headcount ratio at national poverty lines (% of rural population)
-Urban poverty headcount ratio at national poverty lines (% of urban population)
-Poverty headcount ratio at $1.90 a day (2011 PPP) (% of population)
-Poverty headcount ratio at $3.20 a day (2011 PPP) (% of population)
-Poverty headcount ratio at $5.50 a day (2011 PPP) (% of population)
-Poverty gap at $1.90 a day (2011 PPP) (%)
-Poverty gap at $3.20 a day (2011 PPP) (%)
-Poverty gap at $5.50 a day (2011 PPP) (%)
-Poverty gap at national poverty lines (%)

2. Social welfare:
-Coverage of social safety net programs (% of population)
-Coverage of social insurance programs (% of population)


Secondly, we choose to start our analysis in a world level to have a first sight into the data, and then focus on some certain countries with data in relatively abnormal trend. 

Thirdly, we use the data from 1990 to 2015. The number of people living in extreme poverty dropped by more than half between 1990 and 2015. We want to find out the exact trend of how poverty has changed in the word thought out that period. Further more, we try to quantify the correlation between poverty in certain countries and the social welfare system in those countries to have a basic knowledge on how effective the social programs are.


- Data preparation  how did we modify the data to properly use it 
#### Chapter 3 modelling and graphs
-	Define a descriptive model for the relevant data (Visual graphs)
-	Specify a prior distribution
-	Use Bayesian inference to discuss credibility across parameter values
#### Chapter 4 evaluation
- evaluation
-	Discuss research findings to quality of the model, quality of the data. 
-	How could the model be improved  suggest next steps
#### Chapter 5 development
	- Can we predict future behaviour/Guide additional policy
	- How would we communicate our findings to decision makers? 



# START OF THE REPORT

## 1 Introduction to poverty
The world has committed itself to intensive targets for reducing poverty by 2030. The goal to reduce poverty and its related side effects are specified in Sustainable Development Goal (SDG) number one of the total seventeen SDG's set up in the 2030 Agenda for Sustainable Development. All seventeen goals and quantive targets are meant to stimulate action over the next fifteen years in areas of critical importance for humanity and the planet [2]. 

The negative effects of povety are enormous. Poverty and related effects slow down development within whole countries. Certain groups are disproportionately represented and face additional constraints, such as limited access to resources and capabilities in escaping poverty. These include women, persons with disabilities, children, and indigenous peoples.

### 1.1 UN Targets

Eradicating poverty is not about charity, it is an act of justice. It could also unlock an enormous human potential that is to present day have been diminisshed. Still, nearly half of the world’s population lives in poverty. To battle poverty, the UN have constructed several quantative targets that need to be reaced by 2030, see below.

Target 1: By 2030, eradicate extreme poverty for all people everywhere, currently measured as people living on less than $1.25 a day.

Target 2: By 2030, reduce at least by half the proportion of men, women and children of all ages living in poverty in all its dimensions according to national definitions.

Target 3: Implement nationally appropriate social protection systems and measures for all, including floors, and by 2030 achieve substantial coverage of the poor and the vulnerable.

Target 4: By 2030, ensure that all men and women, in particular the poor and the vulnerable, have equal rights to economic resources, as well as access to basic services, ownership and control over land and other forms of property, inheritance, natural resources, appropriate new technology and financial services, including microfinance.

Target 5: By 2030, build the resilience of the poor and those in vulnerable situations and reduce their exposure and vulnerability to climate-related extreme events and other economic, social and environmental shocks and disasters.

These targets have been used to draft our research question, see the next paragraph. 

### 1.2 Goal and research questions

Eradicating poverty is not about charity, it is an act of justice. It could also unlock an enormous human potential that is to present day have been diminisshed. Still, nearly half of the world’s population lives in poverty. To obtain more insight in poverty and its underlying relations causes and effects, the folowwing research question have been constructed, see below.

Main research question: 
How did poverty develop around the world since 1990 to national and absolute dimension and how is this related to the presence of social safety programs?

To answer this research, several sub questions are defined:

-  Question 1: How did poverty for all develop from 1990 – 2015 for national and absolute dimensions?
	- To national definitions  world map
	- To UN definitions 1.90,350,5,50 -> stacked linechart 
	- Distribution for men/women/children stacked bar chart for three groups
	
- Question 2: How is wealth distributed for countries with a relatively high poverty headcount to national poverty lines.
	- GINI index and Lorenz curve for pick five worst from national map
	
- Question 3: What is the effect of social programs on the prevalence of poverty according to national poverty lines?
	- Barchart of absolute and percentage population living under national poverty line
	- Regression with seaborn (python example) x = poverty, y social program

### 1.3 Methodology and Research Goal 
Within this report, we try to gain more insight behind the data of poverty around the world. First, we will clean and analyse the data by selecting the relevant indicators. By means of critical analysis and interpretation of figures and numbers, we try to find relations in the data to obtain useful conclusion for our recommendation to policy makers. Also, relevant primary research on the topic of Zero Hunger and food security will be included to compare our primary research findings.

R-studio is used as a tool to perform the data analysis within this research. R is widely used as a programming language for the visualisation of data. It contains additional packages that are very powerful for data analysis and visualisation purposes. Within our computations in R, several additional packages have been used. For example, Plotly has been used to make our detailed graphics. In Plotly visualisations are displayed in an efficient way and many possibilities to visualize the data are present.


Resources
https://www.globalgoals.org/1-no-poverty
https://sdgcompass.org/sdgs/sdg-1/
https://ourworldindata.org/extreme-poverty


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
